Environmental and geographic variables are effective surrogates for genetic variation in conservation planning

### 2016-11-10 code/rmarkdown/article.Rmd
+ add diagrams to supplementary figures

We created environmental and geographic surrogate variables for each species. To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated the centroid of each grid cell, and extracted their two-dimensional coordinates. To describe the environmental characteristics of each planning unit, we obtained 19 climatic layers ($30^{\prime\prime}$ resolution; [@r33]), projected them and the planning units into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$), and computed planning unit averages for each climatic variable. To reduce dimensionality, for each species, we subject the climatic values associated with their planning units to a principal components analysis (PCA; Table S1). We used the first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components to characterise climatic variation found across the species' distributions inside the study area (capturing `r round(mean(spp.pca.DF[[2]]*100),2)` % $\pm$ `r round(sd(spp.pca.DF[[2]]*100),2)` SD of the total climatic variation; Figure S2). Thus we constructed a two-dimensional geographic space as a potential surrogate for neutral genetic variation, and a `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))`-dimensional environmental space as a potential surrogate for adaptive genetic variation for each species. In these spaces, each planning unit was associated with a single point. Planning units with more similar environmental conditions, or higher spatial proximity in the real world, were associated with points that were closer together in these spaces (respectively).

@@

To investigate the effectiveness of our supposed surrogates, we first needed to identify which of the sampled loci were adaptive. Following best practices, we used two outlier detection methods to achieve this (an individual- and a population-level method; [@r495]). The basic premise underpinning such methods is that neutral loci are expected to exhibit a certain level of variation, and loci that deviate from this expectation are likely to be under selection ([@r480]). The advantage of these methods--in contrast to environmental association analyses--is that they do not use environmental data, which would have introduced circularity into our analysis. Loci identified as outliers in both analyses were treated as adaptive and the remainder were treated as neutral. To minimise false-positives, we omitted loci from both methods where the global frequency of the minor allele was less than 10 % and treated them as neutral.

>>>>

We created environmental and geographic surrogate variables for each species (Figures S59--S561). To describe the geographic location of each planning unit, we projected the grid into an equi-distant coordinate system (Europe Equidistant Conic; $\texttt{ESRI:102031}$), calculated the centroid of each grid cell, and extracted their two-dimensional coordinates. To describe the environmental characteristics of each planning unit, we obtained 19 climatic layers ($30^{\prime\prime}$ resolution; [@r33]), projected them and the planning units into an equal-area coordinate system (Europe Lambert Conformal Conic; $\texttt{ESRI:102014}$), and computed planning unit averages for each climatic variable. To reduce dimensionality, for each species, we subject the climatic values associated with their planning units to a principal components analysis (PCA; Table S1). We used the first `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))` principal components to characterise climatic variation found across the species' distributions inside the study area (capturing `r round(mean(spp.pca.DF[[2]]*100),2)` % $\pm$ `r round(sd(spp.pca.DF[[2]]*100),2)` SD of the total climatic variation; Figure S2). Thus we constructed a two-dimensional geographic space as a potential surrogate for neutral genetic variation, and a `r as.character(as.english(surrogate.params.LST[[MODE]]$number.components))`-dimensional environmental space as a potential surrogate for adaptive genetic variation for each species. In these spaces, each planning unit was associated with a single point. Planning units with more similar environmental conditions, or higher spatial proximity in the real world, were associated with points that were closer together in these spaces (respectively).

@@

To investigate the effectiveness of our supposed surrogates, we quantified the main spatial patterns of adaptive and neutral genetic variation (Figures S59, S562). To do this, we first needed to identify which of the sampled loci were adaptive. Following best practices, we used two outlier detection methods to achieve this (an individual- and a population-level method; [@r495]). The basic premise underpinning such methods is that neutral loci are expected to exhibit a certain level of variation, and loci that deviate from this expectation are likely to be under selection ([@r480]). The advantage of these methods--in contrast to environmental association analyses--is that they do not use environmental data, which would have introduced circularity into our analysis. Loci identified as outliers in both analyses were treated as adaptive and the remainder were treated as neutral. To minimise false-positives, we omitted loci from both methods where the global frequency of the minor allele was less than 10 % and treated them as neutral.

### 2016-11-10 code/rmarkdown/supporting_information.Rmd
+ add diagrams to supplementary figures

>>>>

```{r, include=FALSE}
render_gvis <- function(x) {
  file = tempfile(fileext='.png')
  rsvg::rsvg_png(charToRaw(DiagrammeRsvg::export_svg(x)), file)
  grid::grid.raster(png::readPNG(file))
}
```

```{r, echo=FALSE, results='hide', message=FALSE, fig.cap='Overall data processing workflow. The flowcharts show the methods used to generate the environmental and geographic surrogate spaces, and the adaptive and neutral genetic spaces. Squares indicate data sets. Ellipses indicate analyses. Lines show relationships between data and analysis, and their captions describe how data was processed.'}
render_gvis(grViz("
digraph main {
  graph [layout = dot]

  node [shape = rectangle, style = filled, color = darkslategray, fillcolor = white]

  meirmans_data [label = <Plant data<BR></BR>(Meirmans <I>et al. </I> 2011)>]
  bioclimatic_data [label = <Bioclimatic data<BR></BR>(Hijmans <I> et al. </I> 2005)>]
  sample_location_data [label = 'Sample location data']
  grid_data [label = 'Grid cell data']
  sample_aflp_data [label = 'Sample AFLP data']

  geo_as [label = 'Geographic attribute spaces'] 
  env_as [label = 'Environmental attribute spaces']
  neutral_as [label = 'Neutral genetic attribute spaces']
  adaptive_as [label = 'Adaptive genetic attribute spaces']

  node[shape = ellipse, style = filled, color = darkslategray, fillcolor = white]
  calc_geo_as [label = 'Processing\n(Figure SXXX)'] 
  calc_env_as [label = 'Processing\n(Figure SXXX)']
  calc_genetic_as [label = 'Processing\n(Figure SXXX)']

  edge [color = grey]
  meirmans_data -> {sample_aflp_data grid_data sample_location_data}
  bioclimatic_data -> calc_env_as
  sample_location_data -> {calc_geo_as calc_env_as calc_genetic_as}
  grid_data -> {calc_geo_as calc_env_as calc_genetic_as}
  sample_aflp_data -> calc_genetic_as
  calc_genetic_as -> {neutral_as, adaptive_as}
  calc_geo_as -> geo_as
  calc_env_as -> env_as
  calc_env_as -> calc_env_as [label = ' each species\nseparately']
  calc_geo_as -> calc_geo_as [label = ' each species\nseparately']
  calc_genetic_as -> calc_genetic_as [label = ' each species\nseparately']

}
"))
```

```{r, echo=FALSE, results='hide', message=FALSE, fig.cap='Geographic surrogate processing workflow. Conventions are detailed in Figure S59 caption.'}
render_gvis(grViz("
digraph geo_as {
  graph [layout = dot]

  node [shape = rectangle, style = filled, color = darkslategray, fillcolor = white]

  # input
  sample_location_data [label = 'Sample location data']
  grid_data [label = 'Grid cell data']
  species_name [label = 'Species name']
  {rank = same; sample_location_data; grid_data; species_name}

  # intermediate
  spp_samples [label = 'Locations of samples for species']
  spp_grids [label = 'Grid cells occupied by species']
  spp_proj_data [label = 'Projected grid cells']
  spp_centriod_data [label = 'Centroids of grid cells']

  # output
  geographic_space [label = 'Geographic attribute space']

  edge [color = grey]

  species_name -> spp_samples 
  sample_location_data -> spp_samples [label = ' subset data']
  spp_samples -> spp_grids [label = ' overlay data']
  grid_data -> spp_grids 
  spp_grids -> spp_proj_data [label = ' reproject data to\nequi-distant CRS']
  spp_proj_data -> spp_centriod_data [label = ' compute centroids']
  spp_centriod_data -> geographic_space [label =' extract coordinates and use them\nto represent demand points and\nplanning unit coordinates']

}
"))
```

```{r, fig.height=6, echo=FALSE, results='hide', message=FALSE, fig.cap='Environmental surrogate processing workflow. Conventions are detailed in Figure S59 caption.'}
render_gvis(grViz("
digraph env_as {
  graph [layout = dot]

  node[shape = rectangle, style = filled, color = darkslategray, fillcolor = white]

  # input
  sample_location_data [label = 'Sample location data']
  grid_data [label = 'Grid cell data']
  species_name [label = 'Species name']
  bioclimatic_data [label = <Bioclimatic data<BR></BR>(Hijmans <I> et al. </I> 2005)>]
  {rank = same; sample_location_data; grid_data; species_name; bioclimatic_data}

  # intermediate
  spp_samples [label = 'Locations of samples for species']
  spp_grids [label = 'Grid cells occupied by species']
  spp_proj_data [label = 'Projected grid cells']

  bioclimatic_proj_data [label = 'Projected bioclimatic data']
  clipped_bioclimatic_data [label = 'Bioclimatic data clipped to species distribution']
  pca_scores_data [label = 'Main axes of bioclimatic variation\nacross species distribution']
  grid_pca_data [label = 'Environmental conditions inside each grid cell']

  # output
  environmental_space [label = 'Environmental attribute space']

  # analyses
  node[shape = ellipse, style = filled, color = darkslategray, fillcolor = white]  
  pca [label = 'Principal components analysis']

  edge [color = grey]
  species_name -> spp_samples 
  sample_location_data -> spp_samples [label = ' subset data']
  grid_data -> spp_grids
  bioclimatic_data -> bioclimatic_proj_data [label = ' reproject data to\nequal-area CRS']
  spp_samples -> spp_grids [label = ' overlay data']
  spp_grids -> spp_proj_data [label = ' reproject data to\nequal-area CRS']
  bioclimatic_proj_data -> clipped_bioclimatic_data [label = ' clip data']
  spp_proj_data -> clipped_bioclimatic_data
  clipped_bioclimatic_data -> pca
  pca -> pca_scores_data [label = ' extract scores for\nfirst two components']
  spp_proj_data -> grid_pca_data
  pca_scores_data -> grid_pca_data [label = ' calculate average scores\nfor each grid cell']
  grid_pca_data -> environmental_space [label =' extract scores and use them\nto represent demand points and\nplanning unit coordinates']

}
"))
```

```{r, fig.height=10, echo=FALSE, results='hide', message=FALSE, fig.cap='Genetic data processing workflow. Conventions are detailed in Figure S59 caption.'}
render_gvis(grViz("
digraph genetic_as {
  graph [layout = dot]

  node[shape = rectangle, style = filled, color = darkslategray, fillcolor = white]

  # input
  sample_location_data [label = 'Sample location data']
  sample_aflp_data [label = 'Sample AFLP data']
  grid_data [label = 'Grid cell data']
  species_name [label = 'Species name']
  {rank = same; sample_location_data; grid_data; species_name; sample_aflp_data}

  # intermediate
  spp_aflp_samples [label = 'Species AFLP data']
  spp_location_samples [label = 'Species location data']
  spp_grids [label = 'Grid cells occupied by species']

  population_data [label = 'Samples grouped into populations']
  subset_data [label = 'Samples with reliable population designations']
  rare_alleles [label = 'AFLP data for\nrare alleles']
  remainder_alleles [label = 'AFLP data for\n standard alleles']
  bayescan_adaptive_alleles [label = ' Outlier loci']
  pcadapt_adaptive_alleles [label = ' Outlier loci']
  bayescan_neutral_alleles [label = ' Standard loci']
  pcadapt_neutral_alleles [label = ' Standard loci']
  adaptive_alleles [label = ' AFLP data\nfor adaptive loci']
  neutral_alleles [label = ' AFLP data\nfor neutral loci']
  adaptive_ordinations [label = ' Ordinations showing main\ngradients of variation\nin adaptive loci']
  neutral_ordinations [label = '  Ordinations showing main\ngradients of variation\nin neutral loci']
  grid_adaptive_ordinations [label = ' Typical adaptive genetic characteristics of individuals in grid cells']
  grid_neutral_ordinations [label = ' Typical neutral genetic characteristics of individuals in grid cells']

  # output
  adaptive_genetic_space [label = 'Adaptive genetic space']
  neutral_genetic_space  [label = 'Neutral genetic space']

  # analyses
  node[shape = ellipse, style = filled, color = darkslategray, fillcolor = white]
  bayescan_analysis [label = 'BayeScan analysis']
  pcadapt_analysis [label = 'PCadapt analysis']
  structure_analysis [label = 'Structure analysis']
  clumpp_analysis [label = 'ClumPP analysis']
  neutral_nmds_analysis [label = 'NMDS analyses']
  adaptive_nmds_analysis [label = 'NMDS analyses']
  combine_results [label = 'Combine results from all analyses.\nLoci classified as adaptive in both analyses are treated as adaptive,\notherwise treated as neutral']

  edge [color = grey]
  sample_location_data -> spp_location_samples [label = ' subset data']
  species_name -> spp_location_samples 
  sample_aflp_data -> spp_aflp_samples [label = ' subset data']
  species_name -> spp_aflp_samples
  spp_location_samples -> spp_grids
  grid_data -> spp_grids [label = ' overlay data']
  spp_aflp_samples -> structure_analysis
  structure_analysis -> clumpp_analysis [label = ' post-process\nresults']
  clumpp_analysis -> population_data
  population_data -> subset_data [label = ' omit samples within\nuncertain memberships']
  subset_data -> rare_alleles [label = ' extract rare alleles where\nglobal frequency of minor\nalleles < 10 %']
  subset_data -> remainder_alleles
  remainder_alleles -> bayescan_analysis [label = 'use population\ngroupings']
  remainder_alleles -> pcadapt_analysis [label = 'ignore population\ngroupings']
  bayescan_analysis -> bayescan_adaptive_alleles [label = ' extract']
  pcadapt_analysis -> pcadapt_adaptive_alleles  [label = ' extract']
  bayescan_analysis -> bayescan_neutral_alleles [label = ' extract']
  pcadapt_analysis -> pcadapt_neutral_alleles [label = ' extract']
  bayescan_adaptive_alleles -> combine_results [label = ' treat loci\nas adaptive']
  pcadapt_adaptive_alleles -> combine_results [label = ' treat loci\nas adaptive']
  bayescan_neutral_alleles -> combine_results [label = ' treat loci\nas neutral']
  pcadapt_neutral_alleles -> combine_results [label = ' treat loci\nas neutral']
  rare_alleles -> combine_results [label = ' treat loci\nas neutral']
  combine_results -> adaptive_alleles
  combine_results -> neutral_alleles 
  neutral_alleles  -> neutral_nmds_analysis
  adaptive_alleles -> adaptive_nmds_analysis
  adaptive_nmds_analysis -> adaptive_ordinations [label = ' extract']
  neutral_nmds_analysis  -> neutral_ordinations [label = ' extract']
  spp_grids  -> grid_adaptive_ordinations
  adaptive_ordinations -> grid_adaptive_ordinations [label = ' calculate average values\nin each grid cell']
  spp_grids  -> grid_neutral_ordinations
  neutral_ordinations -> grid_neutral_ordinations [label = ' calculate average values\nin each grid cell']
  grid_adaptive_ordinations -> adaptive_genetic_space [label =' extract values and use them\nto represent demand points and\nplanning unit coordinates']
  grid_neutral_ordinations -> neutral_genetic_space [label =' extract values and use them\nto represent demand points and\nplanning unit coordinates']

}
"))
```
